Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-15 paper/paper.tex
Quick pass

  We partition the input sequence $\sigma$ that arrived during the phase into subsequences $\sigma_I$ and $\sigma_E$, called \emph{internal} and \emph{external}, respectively.
  The subsequence $\sigma_E$ contains requests to pairs that belong to the same component of \TAlg{} at the end of the phase, and $\sigma_E$ denotes the remaining requests issued during this phase.

@@

  \medskip
  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to $\sigma_E$.
  We begin by lower-bounding the number of requests from $\sigma_E$ that $\OPT$ must account (by either serving remotely or migrating), and we denote their number by $x$.
  It holds that $x \geq \TAlg(\sigma_E) - 3\cdot\ell\cdot(\alpha - 1)$, as it received all the requests that \TAlg did, but it could ignore the cost of serving at most $3\cdot\ell$ pairs of nodes, as it might have them collocated in its clusters.
  Each such pair of nodes account for at most $\alpha-1$ requests (the pairs from $\sigma_E$ are unsaturated) in the cost of $\TAlg(\sigma_E)$, hence collocating them would decrease the number by at most $3\cdot\ell\cdot(\alpha - 1)$.

  Now we bound the cost that $\OPT$ pays while serving these $x$ requests.
  Instead of serving all of them remotely, $\OPT$ can perform migrations to decrease its cost.
  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
  (express that between a pair of nodes we accounted $\alpha-1$ requests in $x$)
  Hence, by performing a migration for the cost $\alpha$, $\OPT$ does not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $x$ ($x$ is not a sequence but its length!).
  Finally, for serving $x$, $\OPT$ pays at least $x \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
  This gives us
  \begin{align*}
    \OPT(\sigma_E) + \OPT(M) & \geq (\TAlg(\sigma_E)-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}\\
    &\geq \TAlg(\sigma_E) / 4 - 4\cdot \ell \cdot \alpha
  \end{align*}
  From that, we obtain $\TAlg(\sigma_E) \geq 4\cdot(\OPT(\sigma_E)+\OPT(M)) + 16\ell \cdot \alpha$.

  Now we are ready to bound the competitive ratio.
  Let $E := \OPT(\sigma_E) + \OPT(M)$. Then, using above bounds,
  \begin{equation*}
    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{9\alpha \cdot \ell + \TAlg(\sigma_E)}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{50\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 50 \ell = O(\ell).
  \end{equation*}

@@

  Now we consider the last, unfinished phase.
  Consider the case, where the unfinished phase is also the first one.
  Then, we cannot charge $\OPT$ due to unability to partition the components.
  Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
  We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
  and we follow the analysis regarding the unsaturated requests.
  If the input finished before first $\alpha$ external requests, then \TAlg is optimal.

  If we have at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests
  This way, the competitive ratio increases at most twofold.
\end{proof}

>>>>

  \maciek{Rewrite the paragraph to emphasize in the definition that requests that did not increase the counter are not included}
  We distinguish two disjoint subsequences of input $\sigma$ that arrived during the phase: $\sigma_I$ and $\sigma_E$, called \emph{internal} and \emph{external}, respectively.
  We only account for the requests to nodes that were not collocated by \TAlg at the moment of their arrival, i.e., the requests that increased a counter.
  The subsequence $\sigma_E$ contains requests to pairs that belong to the same component of \TAlg{} at the end of the phase, and $\sigma_E$ denotes the requests to pairs that did not belong to the same component of \TAlg.

@@

  \medskip
  \maciek{The weakest paragraph}
  \maciek{x must be a sequence of requests, not the number; better name for x}
  Before we bound the competitive ratio, we relate the costs of $\TAlg$ and $\OPT$ with respect to $\sigma_E$.
  We begin by lower-bounding the number of requests from $\sigma_E$ that $\OPT$ must account (by either serving remotely or migrating), and we denote their number by $x$.
  It holds that $x \geq \TAlg(\sigma_E) - 3\cdot\ell\cdot(\alpha - 1)$, as it received all the requests that \TAlg did, but it could ignore the cost of serving at most $3\cdot\ell$ pairs of nodes, as it might have them collocated in its clusters.
  Each such pair of nodes account for at most $\alpha-1$ requests (the pairs from $\sigma_E$ are unsaturated) in the cost of $\TAlg(\sigma_E)$, hence collocating them would decrease their number by at most $3\cdot\ell\cdot(\alpha - 1)$.

  Now we bound the cost incurred by $\OPT$ while serving requests from $x$.
  Instead of serving all of them remotely, $\OPT$ can perform migrations to decrease its cost.
  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
  This allows to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
  (express that between a pair of nodes we accounted $\alpha-1$ requests in $x$)
  Hence, by performing a migration for the cost $\alpha$, $\OPT$ does not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $x$ ($x$ is not a sequence but its length!).
  Finally, for serving $x$, $\OPT$ pays at least $x \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
  This gives us
  \begin{equation*}
    \OPT(\sigma_E) + \OPT(M)  \geq (\TAlg(\sigma_E)-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}
    \geq \TAlg(\sigma_E) / 4 - 4\cdot \ell \cdot \alpha.
  \end{equation*}
  From that, we obtain $\TAlg(\sigma_E) \geq 4\cdot(\OPT(\sigma_E)+\OPT(M)) + 16\ell \cdot \alpha$.
  With that, we are ready to bound the competitive ratio.
  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, using above bounds,
  \begin{equation*}
    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{9\alpha \cdot \ell + \TAlg(\sigma_E)}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{50\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 50 \ell = O(\ell).
  \end{equation*}

@@

  Now we consider the last, unfinished phase.
  Consider the case, where the unfinished phase is also the first one.
  Then, we cannot charge $\OPT$ due to unability to partition the components.
  Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
  We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
  and we follow the analysis regarding the unsaturated requests.
  If the input finished before first $\alpha$ external requests, then \TAlg is optimal.
  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests
  This way, the competitive ratio increases at most twofold.
\end{proof}
